%
% file: dedication.tex
% author: V?ctor Bre?a-Medina
% description: Contains the text for thesis dedication
%

\chapter*{Dedication and acknowledgements}
\begin{SingleSpace}
\initial{I} would first like to thank my thesis advisor, Professor Adolfo J. Rumbos of the Mathematics Department of Pomona College. The door to Professor Rumbos's office was always open whenever I ran into a trouble spot or had a question about my research or writing. He consistently allowed this paper to be my own work, but steered me in the right the direction whenever he thought I needed it. \\

I would also like to thank Professors Shahriar Shahriari; Ghassan Sarkis; and Johanna Hardin for how much they've taught me in mathematics. Without their patience, their knowledge, and their passion in teaching, my interests and love for mathematics would not be the same as it is today. \\

I would also like to acknowledge Adam He of Pomona College and Beauttie Kuture of Pomona College. Not only were they helpful second readers of this thesis, but the amount of time spent working together through difficult courses was nothing but invaluable, and it's greatly because of them that this thesis is at the status it is. I see a great future of us working together in the food-truck industry together (just kidding). \\ 

Finally, I must express my most profound gratitude to my beloved \& significant other, Thoa Evelynn Pham, for providing me with unfailing support, continuous encouragement, and love throughout my years of study and through the process of researching and writing this thesis. You have always been by my side physically, mentally, and emotionally, without hesitation for as long as we've been together. None of my accomplishments would not have been possible without you, and I would not be who I am today without you. \\

\begin{otherlanguage*}{vietnamese}
Anh cảm ơn em nhiều, Thoa.
\end{otherlanguage*}

\end{SingleSpace}
\clearpage